---
title: "ASReview FTM Hackathon - Visualization of email domains with VOSViewer"
output: html_notebook
author: Bianca Kramer
date: 2021-11-27
---

## 0. Load packages
```{r message=FALSE, warning=FALSE}

library(tidyverse)
library(readxl)
library(lubridate)
library(tidytext)

```

## 1. Identifying email domains 

## 1a. Read source data 
(n=2509 records)
```{r message=FALSE, warning=FALSE} 

data_preprocessed <- read_excel("../data/preprocessed_data.xlsx", 
                    col_types = c("numeric", "text", "text", 
                                  "text", "date", "text", "text"))

data_preprocessed <- data_preprocessed %>%
  rename(seqID = 1)

#2509 records

#NB column date has inconsistent formatting, but almost all emails have a date in column betterDate

```

## 1b. Retrieve sender and recipient domains from emails
(2109 mail records, 1429 with likely email addresses detected)
```{r message=FALSE, warning=FALSE}

#fiter emails (n=2109 records)
data <- data_preprocessed %>%
  filter(type == "Mail")

#convert body of emails into separate lines (n=291089 lines)
data <- data %>%
  tidytext::unnest_tokens(lines, abstract, token = "lines", drop = FALSE)

#keep lines with (likely) email addresses (n=9628 lines)
#TODO optimize strings used in filter
selection_terms <- c("@", 
                     "to:", "to :", 
                     "from:", "from :",
                     "aan:", "aan :",
                     "van:", "van :")
selection <- str_flatten(selection_terms, "|")

data <- data %>%
  filter(stringr::str_detect(lines, selection))

#check how many original records remain
count_records <- data %>%
  count(id) %>%
  nrow()
#n=1884 (of 2109) records with likely email addresses

#convert remaining lines into separate words (n = 43925 lines)
#NB Tokenizing words removes punctuation (such as @) but keeps periods (.)
#This suits this particular purpose well but good to be aware of
data <- data %>%
  tidytext::unnest_tokens(strings, lines, token = "words", drop = FALSE)

#filter email adress domains xxx.xx  (n= 8732 lines)
data <- data %>%
  filter(stringr::str_detect(strings, "\\.")) 

#check how many original records remain
count_records <- data %>%
  count(id) %>%
  nrow()
#n=1429 (of 1884) records with likely email addresses

rm(selection, selection_terms)

#TODO keep to/from indications to provide directionality
#but many records miss to/from indication, so will loose those

#Note: records with email threads can have multiple to/from email domain strings; records with multiple recipients will also result in multiple email domain strings per record

```

# 1c. Further cleaning of domain strings
(1366 records with 917 unique email domains)

```{r message=FALSE, warning=FALSE}

#read current list of top level domains
tld <- read_tsv("https://data.iana.org/TLD/tlds-alpha-by-domain.txt") %>%
  rename(domain = 1) %>%
  mutate(domain = str_to_lower(domain)) %>%
  pull(domain)

data_clean <- data %>%
  #remove strings with > 1 periods (urls, email prefixes etc)
  filter(str_count(strings, pattern = "\\.") == 1) %>%
  #split domains into prefix and suffix
  separate(strings, c("prefix", "suffix"), sep = "\\.", remove = FALSE) %>%
  #filter on domain suffixes that are valid top level domains
  filter(suffix %in% tld)

#check how many original records remain
count_records <- data_clean %>%
  count(id) %>%
  nrow()
#n=1386 (of 1429) records with likely email addresses

#check how many unique domain strings 
count_strings <- data_clean %>%
  count(strings) %>%
  nrow()
#n=917 (of 7647) unique emai domain strings

rm(tld)

#TODO further clean domain prefixes

```


# 2. Prepare data for VOSviewer  
Create JSON object according to these specs:   https://app.vosviewer.com/docs/file-types/json-file-type/

## 2a. First element in JSON object: node list (called 'items')

```{r message=FALSE, warning=FALSE}

#create df of unique domains with weights as list column
domains <- data_clean %>%
  count(strings) %>%
  distinct() %>%
  arrange(strings) %>%
  mutate(id = row_number()) %>%
  rename(label = strings,
         Documents = n) %>%
  #nest(weights = c(Documents)) %>%
  select(id, label, Documents)

#transpose to list
items <- domains %>%
  purrr::transpose()

#modify list to create a nested list element for weights

#define helper function to modify list elements
modifyList <- function(x){
  x["weights"] <- list(x["Documents"])
  x["Documents"] <- NULL
  return(x)
}

#apply helper function to each list element
items <- map(items, modifyList)



```


## Second element in JSON object: edge list (called 'links')

```{r message=FALSE, warning=FALSE}

a <- items



```
